---
title: "Loading HWSD Data"
author: "Guerzoniansus"
format: pdf
editor: visual
---


```{r}
rm(list = ls())
```

# Install packages

```{r}
# install.packages("sf")
# install.packages("terra")
# install.packages("readxl")
```


# Load packages

```{r}
require(sf)
require(terra)
library(readxl)
library(dplyr)
library(tidyr)
```

# Actions

##  Task 1 : Download the HWSD v2.0 database
 The HWSD is found at the FAO’s Global Agroecological Zoning (GAEZ) page. We do not use the HSWD Viewer on-line application, instead, we download the database and grid for use in R. Two compressed files are provided (Table 1):
 You will not work directly with the MS Access database, rather, the SQLite conversion that is provided with these notes.
 
## Task 2 : Uncompress the raster image HWSD2_raster.zip
 This will create a subdirectory HWSD2_RASTER with four files: the band interleaved image (HWSD2.bil), a small file giving the extent and resolution (HWSD2.stx), the projection information (HWSD2.prj), , and the header (HWSD2.hdr). The latter two are automatically consulted on data import. This raster file is quite large, 1.74 Tb
 
 
```{r}
round(file.size("./HWSD2_RASTER/hwsd2.bil")/1024^3,2)
file.mtime("./HWSD2_RASTER/hwsd2.bil")
```
 
```{r}
hwsd <- rast("./HWSD2_RASTER/HWSD2.bil")
print(hwsd)
```

```{r}
names(hwsd)
```



```{r}
class(hwsd); ncol(hwsd); nrow(hwsd); ncell(hwsd)
res(hwsd); ext(hwsd)
st_crs(hwsd)$proj4string
```

## All techs

```{r}
points <- read.csv("../datasets/I/spam2020V1r0_global_Y_TI.csv")
```


```{r}
head(points)
```


```{r}
points <- points %>% 
  select(c(
    x, y
  ))
```



```{r}
vec_points <- vect(points, geom = c("x", "y"), crs = crs(hwsd))
```

```{r}
values <- terra::extract(hwsd, vec_points)
```

```{r}
values <- cbind(points[, c("x", "y")], values)
```


```{r}
 nrow(values); ncol(values)
```
```{r}
dim(unique(values))
```
```{r}
head(values)
```





```{r}
#| echo: false
#| warning: false


# values[values == 7001] <- NA
layers <- read_excel("./HWSD2_LAYERS/hwsd2_layers.xlsx")
smu <- read_excel("./HWSD2_LAYERS/hwsd2_smu.xlsx")
```

```{r}
layers[layers < 0] <- NA
```



```{r}
layers <- layers %>%
  select(where(is.numeric), LAYER, HWSD2_SMU_ID)
```


```{r}
layers <- layers %>% 
  select(-c(ID, COVERAGE, SEQUENCE, SHARE, ADD_PROP, HWSD1_SMU_ID, ROOT_DEPTH, PHASE1, PHASE2,
    ROOTS, IL, SWR, AWC
  ))
```


```{r}
layers <- layers %>%
  group_by(LAYER, HWSD2_SMU_ID) %>%
  summarise(across(everything(), ~ mean(.x, na.rm = TRUE), .names = "{col}"))
```


```{r}
layers[layers == "NaN"] <- NA

layers <- layers %>% 
  select(-c(TEXTURE_USDA))
```


```{r}
layers_wide <- layers %>%
  pivot_wider(
    id_cols = c(HWSD2_SMU_ID), 
    names_from = LAYER, 
    values_from = all_of(names(layers)[!names(layers) %in% c("HWSD2_SMU_ID", "LAYER")]))
```




```{r}
#smu <- smu %>% 
  #select(c(
    #ID, HWSD2_SMU_ID, COVERAGE, SHARE, WRB2_CODE, TEXTURE_USDA, REF_BULK_DENSITY, BULK_DENSITY,
    #ROOT_DEPTH, AWC
  #))
```

```{r}
crop.layers <- merge(values, layers_wide, by.x = "HWSD2", by.y = "HWSD2_SMU_ID", all.x=TRUE)
```


```{r}
write.csv(crop.layers, file = "data/hwsd2.csv", row.names = FALSE)
```


```{r}
write.csv(layers_wide, file = "data/layers.csv", row.names = FALSE)
```











